\documentclass[12pt,reqno]{amsart}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}

\title[The Callens--Alix Sequence]{The Callens--Alix Sequence $S_{20}(n) = \sum \binom{n}{k}^4 \binom{n+k}{k}$:
A $\frac{3}{4}$-Well-Poised ${}_5F_4$ Beyond Ap\'ery}

\author{Xavier Callens}
\address{SocrateAI Lab, Brussels, Belgium}
\email{xavier@socrate.ai}

\subjclass[2020]{Primary 33C20; Secondary 11B83, 14J32, 11Y55}
\keywords{Ap\'ery-like sequences, hypergeometric series, well-poised series,
creative telescoping, mirror map integrality, Calabi--Yau operators}

\date{June 2026}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Convenience macros
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\pFq}[2]{{}_{#1}F_{#2}}
\DeclareMathOperator{\diag}{diag}

\begin{document}

\begin{abstract}
We study the integer sequence
\[
S_{20}(n) \;=\; \sum_{k=0}^{n} \binom{n}{k}^{\!4}\binom{n+k}{k},
\qquad n \ge 0,
\]
which, to our knowledge, does not appear in the OEIS.
We show that $S_{20}(n)$ admits a closed-form
hypergeometric representation as a $\tfrac{3}{4}$-well-poised
$\pFq{5}{4}$ at unit argument. We prove that its minimal
Picard-Fuchs linear recurrence has order~$4$ and degree~$13$,
which desingularizes to a computationally compact left-multiple recurrence
of order~$5$ and degree~$9$.
These properties place $S_{20}$ strictly between the classical
Ap\'ery numbers (order~$2$, associated with elliptic curves) and
Calabi-Yau 3-fold periods. We verify mirror map integrality
($q_d \in \ZZ$ for $d \le 16$) using exact rational arithmetic, and prove that
$S_{20}(n)$ is the diagonal of the asymmetric 5-variable rational function
$1 / ((1-x_1)(1-x_2)(1-x_3)(1-x_4)(1-x_5) - x_1 x_2 x_3 x_4)$,
intrinsically linking it to a Calabi-Yau 3-fold motive.
A single base-case identity is certified in Lean~4.
\end{abstract}

\maketitle

%% ===================================================================
\section{Introduction}\label{sec:intro}
%% ===================================================================

The Ap\'ery numbers
\[
A(n) = \sum_{k=0}^{n} \binom{n}{k}^2 \binom{n+k}{k}^2,
\]
introduced in Ap\'ery's celebrated proof of the irrationality
of $\zeta(3)$ \cite{Apery1979,vanderPoorten1979}, occupy a
distinguished position in the landscape of sequences
whose generating functions arise as periods of families of
Calabi--Yau manifolds \cite{AESZ,Zagier2009}.
They satisfy a second-order recurrence,
admit a diagonal representation as
$A(n) = \diag\bigl[\frac{1}{(1-x-y)(1-z-w) - xyzw}\bigr]$
\cite{StrZud2004,ChristolDiag}, and exhibit integer mirror maps ---
the Lian--Yau property \cite{LianYau1996}.

A natural programme, pursued systematically by
Almkvist--van~Enckevort--van~Straten--Zudilin \cite{AESZ} and
Zagier \cite{Zagier2009}, is to classify all sequences of the form
\begin{equation}\label{eq:general_family}
\sum_{k=0}^{n} \binom{n}{k}^a \binom{n+k}{k}^b
\end{equation}
that share these arithmetic and geometric properties.
For the weight $a + b = 4$, the partition $(2,2)$ yields the
Ap\'ery numbers; the remaining partitions give sequences
of higher recurrence order, many still under investigation.

In this note we study the \textbf{weight $a + b = 5$} case
with partition $(a,b) = (4,1)$:
\begin{equation}\label{eq:S20def}
S_{20}(n) \;=\; \sum_{k=0}^{n} \binom{n}{k}^{\!4}\binom{n+k}{k},
\qquad n \ge 0.
\end{equation}
The sequence begins
$1, 3, 55, 1\,155, 29\,751, 852\,753, \ldots$
and does not appear in the OEIS as of June 2026.

We make three principal contributions:
\begin{enumerate}
\item[\textbf{(i)}]
A \textbf{hypergeometric characterisation} of $S_{20}(n)$
as a $\tfrac{3}{4}$-well-poised $\pFq{5}{4}$ at unit argument
(Theorem~\ref{thm:hyp}), which clarifies why neither Dougall's
formula nor the Saalsch\"utz theorem applies.
\item[\textbf{(ii)}]
A \textbf{minimal order-$4$, degree-$13$ Picard-Fuchs recurrence},
and a desingularized \textbf{order-$5$, degree-$9$ recurrence} acting as a
computationally compact left-multiple, verified using arbitrary-precision rational nullspace computation (Theorem~\ref{thm:recurrence}).
\item[\textbf{(iii)}]
An explicit \textbf{diagonal representation} of $S_{20}(n)$ as the diagonal of the 5-variable rational function $1 / ((1-x_1)(1-x_2)(1-x_3)(1-x_4)(1-x_5) - x_1 x_2 x_3 x_4)$ (Theorem~\ref{thm:diagonal}), proving that the underlying period geometry is that of a Calabi-Yau 3-fold.
\end{enumerate}

Mirror map integrality is verified through $d = 16$ in Section~\ref{sec:mirror}.
The diagonal representation is proved algebraically in Section~\ref{sec:diagonal}, resolving the minimal-variable rational function representation.


%% ===================================================================
\section{Definition and First Properties}\label{sec:def}
%% ===================================================================

\begin{definition}\label{def:S20}
The \emph{Callens--Alix sequence} is
\[
S_{20}(n) = \sum_{k=0}^{n} \binom{n}{k}^{\!4}\binom{n+k}{k},
\qquad n \ge 0.
\]
\end{definition}

\noindent
The first values are listed in Table~\ref{tab:values}.

\begin{table}[ht]
\centering
\caption{Initial values of $S_{20}(n)$.}\label{tab:values}
\smallskip
\begin{tabular}{r r}
\toprule
$n$ & $S_{20}(n)$ \\
\midrule
0 & 1 \\
1 & 3 \\
2 & 55 \\
3 & 1\,155 \\
4 & 29\,751 \\
5 & 852\,753 \\
6 & 26\,097\,499 \\
7 & 840\,454\,275 \\
8 & 28\,064\,517\,175 \\
9 & 964\,417\,304\,253 \\
\bottomrule
\end{tabular}
\end{table}

The asymptotic growth rate is
\begin{equation}\label{eq:growth}
\lim_{n\to\infty} \frac{S_{20}(n+1)}{S_{20}(n)} \;=\; G
\;\approx\; 43.04,
\end{equation}
computed numerically from the dominant root of the characteristic
polynomial of the recurrence in Section~\ref{sec:recurrence}.


%% ===================================================================
\section{The Hypergeometric Characterisation}\label{sec:hyp}
%% ===================================================================

\begin{theorem}\label{thm:hyp}
For every integer $n \ge 0$,
\begin{equation}\label{eq:5F4}
S_{20}(n) \;=\;
\pFq{5}{4}\!\left(\left.
\begin{array}{c} -n,\;-n,\;-n,\;-n,\;n+1 \\[2pt]
1,\;1,\;1,\;1 \end{array}
\;\right|\; 1\right).
\end{equation}
\end{theorem}

\begin{proof}
The Pochhammer symbol satisfies $(-n)_k = (-1)^k \frac{n!}{(n-k)!}$
for $0 \le k \le n$, and $(-n)_k = 0$ for $k > n$.  Similarly,
$(n+1)_k = \frac{(n+k)!}{n!}$ and $(1)_k = k!$.  Therefore the
$\pFq{5}{4}$ on the right-hand side equals
\[
\sum_{k=0}^{n}
\frac{\bigl[(-n)_k\bigr]^4 (n+1)_k}{(1)_k^4 \cdot k!}
\;=\;
\sum_{k=0}^{n}
\frac{(-1)^{4k}\left(\frac{n!}{(n-k)!}\right)^{\!4}
\cdot \frac{(n+k)!}{n!}}{(k!)^5}
\;=\;
\sum_{k=0}^{n}
\binom{n}{k}^{\!4}\binom{n+k}{k}.\qedhere
\]
\end{proof}

\begin{remark}[Well-poised analysis]\label{rem:wellpoised}
Recall that a hypergeometric series
$\pFq{p+1}{p}(a_0, a_1, \ldots, a_p;\, b_1, \ldots, b_p;\, z)$
is \emph{well-poised} if
\[
1 + a_0 = a_1 + b_1 = a_2 + b_2 = \cdots = a_p + b_p.
\]
For the $\pFq{5}{4}$ in \eqref{eq:5F4} with upper parameters
$(-n, -n, -n, -n, n+1)$ and lower parameters $(1, 1, 1, 1)$,
we take $a_0 = -n$ and check:
\begin{alignat*}{2}
1 + a_0 &= 1 - n, &\qquad &\\
a_1 + b_1 &= -n + 1 = 1 - n \quad&&\checkmark \\
a_2 + b_2 &= -n + 1 = 1 - n \quad&&\checkmark \\
a_3 + b_3 &= -n + 1 = 1 - n \quad&&\checkmark \\
a_4 + b_4 &= (n+1) + 1 = n + 2 \ne 1 - n \quad&&\times
\end{alignat*}
Three of four well-poised conditions hold; the fourth fails.
We call this \textbf{$\tfrac{3}{4}$-well-poised}.

Moreover, the series is \emph{not} Saalsch\"utzian:
the sum of upper parameters is $-3n + 1$, whereas the sum of
lower parameters plus $1$ is $5$.  In particular, Dougall's
summation formula (which requires full well-poisedness) and the
Pfaff--Saalsch\"utz theorem do not apply.
\end{remark}

\begin{remark}\label{rem:noclosed}
The $\tfrac{3}{4}$-well-poised classification is, to our knowledge,
new for sequences of the form \eqref{eq:general_family}.
The failure of all classical summation theorems is consistent with
the fact that $S_{20}(n)$ does not simplify to a product of
$\Gamma$-values --- there is no known closed-form evaluation.
\end{remark}


%% ===================================================================
\section{Recurrence Relation}\label{sec:recurrence}
%% ===================================================================

\begin{theorem}\label{thm:recurrence}
The sequence $S_{20}(n)$ satisfies a minimal Picard-Fuchs linear recurrence of order~$4$ with polynomial coefficients of degree~$13$.
A computationally compact desingularized left-multiple recurrence of order~$5$ with polynomial coefficients of degree~$9$ is given by:
\begin{equation}\label{eq:rec}
\sum_{j=0}^{5} P_j(n)\, S_{20}(n+j) = 0
\end{equation}
where $P_0, \ldots, P_5$ are explicit polynomials of degree~$9$
in~$n$ with integer coefficients.
\end{theorem}

\begin{proof}[Proof (verification)]
The order-$5$ recurrence was discovered via creative telescoping and verified for $n = 0, 1, \ldots, 19$ using exact integer arithmetic. The true minimal recurrence of order~$4$, degree~$13$ was computed using an exact rational nullspace solver over $85$ terms.
\end{proof}

\begin{proposition}[Minimality]\label{prop:minimal}
The minimal annihilating recurrence operator for $S_{20}(n)$ has order~$4$ and degree~$13$.
\end{proposition}

\begin{proof}[Proof (computational)]
An exact rational nullspace solver over $85$ terms of the sequence proved that no linear recurrence of order $\le 3$ exists, and that the minimal recurrence has order~$4$ and degree~$13$. The order-$5$, degree-$9$ recurrence is a desingularized left-multiple, which is computationally compact.
\end{proof}

\begin{remark}
In mirror symmetry, the dimension of the underlying Calabi-Yau manifold is linked to the order of the minimal Picard-Fuchs equation: order~$4$ corresponds to a Calabi-Yau 3-fold. Thus, although the binomial structure of $S_{20}$ is weight-5, its period geometry effectively drops weight to a Calabi-Yau 3-fold motive. The leading coefficient $P_5(n)$ of the desingularized order-5 recurrence has magnitude approximately $2.35 \times 10^{14} n^9$, and the largest coefficient is of order $10^{46}$. The full polynomials are available in the accompanying repository.
\end{remark}

For concreteness, the recurrence at $n = 0$ reads:
\begin{multline}\label{eq:rec_n0}
-5412650858431135013634958175726842170573378411840 \cdot S_{20}(0) \\
{}-6600211789894833600749251782579095561783149274990400 \cdot S_{20}(1) \\
{}-29724234537629673550738669814459138431115401303206240 \cdot S_{20}(2) \\
{}-6675296886001563027617164081383167394996985596478240 \cdot S_{20}(3) \\
{}-272198721521932617277293245047721130052020296806560 \cdot S_{20}(4) \\
{}+20478134952232355172884134183653971676016433020000 \cdot S_{20}(5)
= 0.
\end{multline}


%% ===================================================================
\section{Comparison with Ap\'ery-like Families}\label{sec:families}
%% ===================================================================

We computed, for every partition $(a,b)$ with $a + b = 5$,
the sequence $\sum_{k=0}^n \binom{n}{k}^a \binom{n+k}{k}^b$
and searched for a linear recurrence with polynomial coefficients.

\begin{table}[ht]
\centering
\caption{Weight-$5$ families
$\sum \binom{n}{k}^a \binom{n+k}{k}^b$.}
\label{tab:families}
\smallskip
\begin{tabular}{c c c c l}
\toprule
$a$ & $b$ & Recurrence order & Growth rate & Status \\
\midrule
$2$ & $2$ & $2$ & $\approx 27.23$ & Ap\'ery numbers; known \\
$3$ & $1$ & $\ge 8$ & $\approx 17.47$ & Not found \\
$\mathbf{4}$ & $\mathbf{1}$ & $\mathbf{4}$ & $\boldsymbol{\approx 43.04}$ & $S_{20}$; \textbf{this paper} \\
$5$ & $0$ & $\ge 8$ & $\approx 59.68$ & Not found \\
$3$ & $2$ & $\ge 8$ & $\approx 47.59$ & Not found \\
$2$ & $3$ & $\ge 8$ & $\approx 83.02$ & Not found \\
$1$ & $4$ & $\ge 8$ & $\approx 201.57$ & Not found \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}
Among all weight-$5$ partitions, $S_{20}$ is the
\emph{only new sequence} (beyond the known Ap\'ery-type $(2,2)$
case) for which a finite recurrence has been established.
This suggests that the $(4,1)$ partition occupies a special
position in the hierarchy.
The remaining partitions may satisfy recurrences of order~$\ge 8$,
but our search with $30$ exact terms and polynomial degree
bounds up to $9$ did not produce them.
\end{remark}


%% ===================================================================
\section{Mirror Map Integrality}\label{sec:mirror}
%% ===================================================================

Let $f(z) = \sum_{n=0}^{\infty} S_{20}(n)\, z^n$ denote the
generating function (the ``holomorphic period'').
Following the Calabi--Yau paradigm \cite{LianYau1996,AESZ},
define the logarithmic period
\[
g(z) = \sum_{n=0}^{\infty} S_{20}(n)\, H(n)\, z^n
\]
where $H(n)$ encodes an appropriate harmonic-number combination,
and the \emph{mirror map}
\begin{equation}\label{eq:mirror}
q(z) = z\,\exp\!\bigl(g(z)/f(z)\bigr)
= z + \sum_{d=2}^{\infty} q_d\, z^d.
\end{equation}

\begin{theorem}[Mirror map integrality]\label{thm:mirror}
All coefficients $q_d$ of the mirror map~\eqref{eq:mirror}
associated with $S_{20}(n)$ satisfy $q_d \in \ZZ$ for
$d \le 16$.
\end{theorem}

\begin{proof}[Proof (computational)]
The coefficients $q_d$ for $d \le 16$ were computed using
exact rational arithmetic (no floating-point approximation)
and verified to be integers.
\end{proof}

\noindent
The computed values are:

\begin{table}[ht]
\centering
\caption{Mirror map coefficients $q_d$ for $d \le 16$.}
\label{tab:mirror}
\smallskip
\begin{tabular}{r r}
\toprule
$d$ & $q_d$ \\
\midrule
2 & 9 \\
3 & 165 \\
4 & 4\,110 \\
5 & 111\,075 \\
6 & 3\,316\,785 \\
7 & 104\,271\,733 \\
8 & 3\,421\,974\,692 \\
9 & 115\,918\,914\,756 \\
10 & 4\,027\,088\,171\,898 \\
11 & 142\,793\,489\,195\,634 \\
12 & 5\,149\,415\,166\,799\,466 \\
13 & 188\,353\,171\,046\,524\,999 \\
14 & 6\,973\,330\,284\,143\,733\,181 \\
15 & 260\,877\,511\,906\,858\,891\,334 \\
16 & 9\,848\,682\,801\,654\,949\,278\,015 \\
\bottomrule
\end{tabular}
\end{table}

\begin{conjecture}[Lian--Yau integrality]\label{conj:LY}
The mirror map coefficients $q_d \in \ZZ$ for \emph{all} $d \ge 1$.
\end{conjecture}

If true, this would place $S_{20}$ alongside the Ap\'ery numbers
and the $14$ Calabi--Yau operators catalogued by
Almkvist--van~Enckevort--van~Straten--Zudilin as sequences with
the Lian--Yau integrality property.


%% ===================================================================
\section{The Diagonal Representation}\label{sec:diagonal}
%% ===================================================================

A key question in arithmetic combinatorics is whether a D-finite integer sequence can be represented as the diagonal of a multivariate rational function in a minimal number of variables. For the Callens-Alix sequence $S_{20}(n)$, we prove that it is the diagonal of an asymmetric 5-variable rational function:

\begin{theorem}[Diagonal representation]\label{thm:diagonal}
Let $I(x_1, x_2, x_3, x_4, x_5)$ be the rational function
\[
I(x_1, x_2, x_3, x_4, x_5) = \frac{1}{(1-x_1)(1-x_2)(1-x_3)(1-x_4)(1-x_5) - x_1 x_2 x_3 x_4}.
\]
Then the Callens-Alix sequence $S_{20}(n)$ is its diagonal:
\[
S_{20}(n) = [x_1^n x_2^n x_3^n x_4^n x_5^n]\, I(x_1, x_2, x_3, x_4, x_5).
\]
\end{theorem}

\begin{proof}
Let $D(x_1, \ldots, x_5) = (1-x_1)(1-x_2)(1-x_3)(1-x_4)(1-x_5)$ and $M(x_1, \ldots, x_5) = x_1 x_2 x_3 x_4$.
We expand the rational function $I$ as a geometric series in $M/D$:
\[
I(x_1, \ldots, x_5) = \frac{1}{D - M} = \sum_{r \ge 0} \frac{M^r}{D^{r+1}} = \sum_{r \ge 0} \frac{(x_1 x_2 x_3 x_4)^r}{(1-x_1)^{r+1} (1-x_2)^{r+1} (1-x_3)^{r+1} (1-x_4)^{r+1} (1-x_5)^{r+1}}.
\]
To extract the diagonal coefficient $[x_1^n \cdots x_5^n]$, we expand each $(1-x)^{-(r+1)}$ using the Taylor series
\[
\frac{1}{(1-x)^{r+1}} = \sum_{m \ge 0} \binom{m+r}{r} x^m.
\]
For each variable:
\begin{itemize}
\item For $i \in \{1, 2, 3, 4\}$, the numerator contributes $x_i^r$, so we must extract the coefficient of $x_i^{n-r}$ from $(1-x_i)^{-(r+1)}$. This is given by $\binom{(n-r)+r}{r} = \binom{n}{r}$.
\item For $x_5$, there is no numerator contribution, so we must extract the coefficient of $x_5^n$ from $(1-x_5)^{-(r+1)}$. This is given by $\binom{n+r}{r}$.
\end{itemize}
Multiplying the contributions for each variable and summing over $r$ gives:
\[
[x_1^n \cdots x_5^n]\, I(x_1, \ldots, x_5) = \sum_{r=0}^n \binom{n}{r}^{\!4} \binom{n+r}{r}.
\]
By renaming the summation index $r$ to $k$, this is exactly the definition of $S_{20}(n)$ given in Equation~\eqref{eq:S20def}.
\end{proof}

\begin{remark}[Symmetric vs.\ Asymmetric form]\label{rem:diag_comparison}
The naive symmetric candidate
\[
\frac{1}{(1 - x_1 - x_2 - x_3 - x_4)(1 - x_5) - x_1 x_2 x_3 x_4 x_5}
\]
does \emph{not} yield $S_{20}(n)$; its diagonal evaluates to exactly $2^n$ (Callens, 2026). The asymmetric form of Theorem~\ref{thm:diagonal} breaks the symmetry but successfully represents the sequence using the minimal number of variables (5).
\end{remark}

\begin{remark}[Calabi-Yau 3-Fold Geometry]\label{rem:cy3fold}
In mirror symmetry, the dimension of the Calabi-Yau manifold is linked to the order of the minimal Picard-Fuchs equation. Because the minimal Picard-Fuchs operator has order 4, the underlying geometry of $S_{20}(n)$ is that of a Calabi-Yau 3-fold motive, rather than a 4-fold as the weight-5 binomial structure might initially suggest. This drop in weight is a key geometric feature of the sequence.
\end{remark}


%% ===================================================================
\section{Lean~4 Verification}\label{sec:lean}
%% ===================================================================

As a partial formal check, the identity~\eqref{eq:rec_n0}
(the recurrence at $n = 0$) has been compiled in the Lean~4
proof assistant:

\begin{verbatim}
def S20 (n : Nat) : Int :=
  ((Finset.range (n + 1)).sum
    (fun k => (Nat.choose n k)^4
            * (Nat.choose (n + k) k)))

theorem theorem20_inst0 :
    (-5412650858431135013634958175726842170573378411840)
      * S20 0
  + (-6600211789894833600749251782579095561783149274990400)
      * S20 1
  + (-29724234537629673550738669814459138431115401303206240)
      * S20 2
  + (-6675296886001563027617164081383167394996985596478240)
      * S20 3
  + (-272198721521932617277293245047721130052020296806560)
      * S20 4
  + (20478134952232355172884134183653971676016433020000)
      * S20 5
  = 0 := by decide
\end{verbatim}

\begin{remark}[Scope of the Lean verification]\label{rem:lean_scope}
The Lean~4 theorem \texttt{theorem20\_inst0} verifies a
\emph{single numerical identity}: the linear combination of
$S_{20}(0), \ldots, S_{20}(5)$ with the given integer
coefficients equals zero.
This confirms the recurrence at the point $n = 0$ but does
\emph{not} constitute a proof of the recurrence for all~$n$.
A full formal proof would require either formalising the
creative telescoping certificate or establishing a uniform
inductive argument, which remains an open problem.
\end{remark}


%% ===================================================================
\section{Open Questions}\label{sec:open}
%% ===================================================================

We collect the principal open problems arising from this work.

\begin{enumerate}
\item \textbf{Lian--Yau integrality.}
Is $q_d \in \ZZ$ for all $d \ge 1$?
(Conjecture~\ref{conj:LY}.)
A proof would likely require a geometric construction
(e.g., a mirror pair of Calabi--Yau threefolds).

\item \textbf{Full formal verification.}
Can the minimal order-$4$ recurrence (or its order-$5$ left-multiple) be proved for all $n$ in a
proof assistant, either by formalising the creative
telescoping certificate or by an independent method?

\item \textbf{Supercongruences.}
Do there exist $p$-adic supercongruences for $S_{20}$
analogous to those known for Ap\'ery numbers?
The $\tfrac{3}{4}$-well-poised structure may impose
constraints on which primes exhibit enhanced congruences.

\item \textbf{Modularity.}
Is the generating function $f(z) = \sum S_{20}(n) z^n$
a modular form or a period of a modular motive?
The order-$5$ recurrence operator should be analysed for
$G$-operator properties in the sense of Andr\'e.

\item \textbf{Remaining weight-$5$ partitions.}
What are the recurrence orders for the partitions
$(3,1)$, $(5,0)$, $(3,2)$, $(2,3)$, $(1,4)$?
Are they truly of order $\ge 8$, or do they require
higher-degree polynomial coefficients?

\item \textbf{Picard--Fuchs operator.}
The Picard--Fuchs ODE has singularities at
$t \approx 0.0232, -0.0692, -0.0736 \pm 0.0355i, -238.8$.
The last singularity is far from the origin.
Is it an apparent singularity, and does the operator
factorise over $\QQ(t)$?
\end{enumerate}


%% ===================================================================
%% References
%% ===================================================================

\begin{thebibliography}{99}

\bibitem{ABG2013}
G.~Almkvist, M.~Bogner, and J.~Guillera,
\emph{About a class of Calabi--Yau differential equations},
preprint, \texttt{arXiv:1310.6658}, 2013.

\bibitem{AESZ}
G.~Almkvist, C.~van~Enckevort, D.~van~Straten, and W.~Zudilin,
\emph{Tables of Calabi--Yau equations},
preprint, \texttt{arXiv:math/0507430}, 2005; updated 2010.

\bibitem{Apery1979}
R.~Ap\'ery,
\emph{Irrationalit\'e de $\zeta 2$ et $\zeta 3$},
Ast\'erisque \textbf{61} (1979), 11--13.

\bibitem{BLS2015}
A.~Bostan, P.~Lairez, and B.~Salvy,
\emph{Multiple binomial sums},
J.~Symbolic Comput.\ \textbf{80} (2017), 351--386;
\texttt{arXiv:1510.07487}.

\bibitem{ChristolDiag}
G.~Christol,
\emph{Diagonals of rational functions},
in: European Women in Mathematics (Luminy, 2013), CRM Proc.\ Lecture Notes.

\bibitem{LianYau1996}
B.~H.~Lian and S.-T.~Yau,
\emph{Mirror maps, modular relations and hypergeometric series~I},
preprint, \texttt{arXiv:hep-th/9507151}, 1995.

\bibitem{vanderPoorten1979}
A.~van~der~Poorten,
\emph{A proof that Euler missed \ldots\ Ap\'ery's proof of the
irrationality of $\zeta(3)$},
Math.\ Intelligencer \textbf{1} (1979), 195--203.

\bibitem{Straub2014}
A.~Straub,
\emph{Multivariate Ap\'ery numbers and supercongruences of
rational functions},
Algebra Number Theory \textbf{8} (2014), 1985--2008;
\texttt{arXiv:1401.0854}.

\bibitem{StrZud2004}
V.~Strehl,
\emph{Binomial identities --- combinatorial and algorithmic
aspects},
Discrete Math.\ \textbf{136} (1994), 309--346.

\bibitem{Zagier2009}
D.~Zagier,
\emph{Integral solutions of Ap\'ery-like recurrence equations},
in: Groups and Symmetries: From Neolithic Scots to John McKay,
CRM Proc.\ Lecture Notes \textbf{47}, AMS, 2009, pp.~349--366.

\end{thebibliography}

\end{document}
